---
title: "Naples Reef Purple Urchins"
author: "Alissa Patterson"
date: "10/29/2021"
output: html_document
---

## Overview
### This code wrangles, analyses, and visualizes data of Naples Reef observations for purple urchins observed in 2010.


Data: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

## Setup
## Attach packages

```{r}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(knitr)
library(kableExtra)
```

# Read in the data

```{r}
urchins <- read_csv(here("data","urchins.csv"))
urchins <- clean_names(urchins)
```

# Data wrangling

Select only the Naples Reef observations for purple urchins in 2010. 

```{r}
naples_urchins <- urchins %>% 
  filter(site == "NAPL" &  common_name == "Purple Urchin") %>%
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>%
  filter(year == 2010)
```


# Data exploration

```{r}
ggplot(naples_urchins, aes(size)) +
  geom_histogram(bins=10) +
  facet_grid(~ treatment) +
  theme_minimal() +
  labs( x = "Size (cm)",
        caption = "Histogram of Purple Urchin Sizes (cm) by Treatment") +
  theme(plot.caption = element_text(hjust = 0))

ggplot(naples_urchins, aes(sample=size)) +
  geom_qq() +
  stat_qq_line() +
  facet_grid(~ treatment) +
  theme_minimal() +
  labs( y = "Size (cm)",
        x = "Theoretical",
        caption = "QQ-plot of purple urchin sizes per treatment") +
  theme(plot.caption = element_text(hjust = 0))
```

The data distribution is not normal. In the histograms, the distributions for both treatments are skewed to the right. In the QQ-plot, the data does not follow the normal distribution line. Although the distributions of the treatment and control are very similar, the annual treatment resulted in a high count of urchins sized 3 cm compared to the control. Overall, removing kelp does not appear to have a large effect on the purple urchins.  


# Descriptive statistics

Create a finalized summary table that contains some statistics for annual and control samples.

```{r}
naples_urchins_summary <- naples_urchins %>% 
  group_by(treatment) %>% 
  summarize(mean=mean(size), median=median(size), sd=sd(size), n=n())

kable(naples_urchins_summary,
      digits = 2,
      col.names = c("Treatment",
                    "Mean (cm)",
                    "Median (cm)",
                    "Standar deviation (cm)",
                    "Sample size"),
      caption = "Summary statistics by treatment for Naples Reef's purple urchins in 2010 samples") %>% 
  kable_styling(full_width = FALSE)
```





